---
title: "Middle School Network"
author: "Chong Zhang"
date: 5/22/2023
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```


```{r, warning=FALSE }
library(igraph);library(netplot);
library(data.table);require(dplyr)
library(network);library(ergm);library(intergraph)


students     <- fread("data/pone.0153690.s001.csv") #678
interactions <- fread("data/pone.0153690.s003.csv") #10781
```

## Data cleaning 
Remove interactions involving missing ids and node attributes
```{r }
students <- students[!is.na(id)] #639 - remove missing ids
students=students[gender!=99 & lunch!=99, ] # remove missing attributes
interactions <- interactions[!is.na(id) & !is.na(contactId), .(id, contactId)] %>% unique # keep unique interactions.
head(interactions) 
dim(interactions)
```


```{r }
setdiff(c(interactions$id, interactions$contactId), students$id)
```
Some actors in the interaction files were not found in the student file (or removed) including '99' - not identified based on information provided, and we will removed them.

```{r }
# Checking which connections are not OK
ids <- sort(unique(students$id)) 
interactions <- interactions[(id %in% ids) & (contactId %in% ids)] 
dim(interactions)
```

We will also remove self interactions (id=contactId)
```{r }
interactions=interactions[id!=contactId] 
dim(interactions)
```

Sometimes we see interactions recorded in both directions:
```{r , echo=FALSE}
interactions[id==2004 & contactId==2028]
interactions[id==2028 & contactId==2004] # this is the same edge duplicated
```
Since the network is undirected these are considered the same. We will keep only one. 
find out how many such duplicates are there
```{r }
interactions.r=interactions[,.(contactId, id)] # switch the contactId and id role 
names(interactions.r)=c('id','contactId') 
interactions.r$rev=1 # indicator for 'reversed' interaction

int.m=merge(interactions, interactions.r, by=c('id', 'contactId'), all.x=TRUE) 
edges=int.m[is.na(rev) | id<contactId] # keep only interactions in the first direction (id < contactId)
dim(edges)
```

## Now build the network
We don't want isolates, so ids not forming any edges are removed.
```{R }
nodes=students[id %in% c(edges$id, edges$contactId)] 

net <- graph_from_data_frame(
  d = edges,
  directed = FALSE, vertices = as.data.frame(nodes)
)
```

## Network statistics 
Edge count, mean degree, number of isolates, density, modularity...
```{r}
ecount(net) #edges

hist(degree(net), breaks = 50) # degree
mean(degree(net)) 

edge_density(net)  # density

groups=cluster_edge_betweenness(net) #clustering
length(groups) # now only 16!
modularity(groups) # 0.407
```
Plot the network
```{r}
set.seed(1234)
sg1=net
sg=asNetwork(sg1)

col=c("8"="seagreen", "7"="blue")
cols=col[as.character(sg %v% "grade")]

nplot_base(sg1,  vertex.color=cols, edge.arrow.size=0 )
legend("bottomright", legend=c("7", "8"), fill=col,  title="Grade")
```


Plot just the first 150 vertices and ties between them (some did not form ties within this subgroup, but had ties with other students)
```{r}
set.seed(123)
sg1=induced_subgraph(net, 1:150)
#sg1=net
sg=asNetwork(sg1)

col=c("8"="seagreen", "7"="blue")
shap=c("0"=15, "1"=3)
cols=col[as.character(sg %v% "grade")]
shapes=shap[as.character(sg %v% 'gender')]

#sg %v% "color" = cols
#sg %v% "shape" =shapes

class(sg1)

nplot_base(sg1,  vertex.color=cols, vertex.nsides = shapes,  edge.curvature=0, edge.arrow.size=0 )
legend("bottomright", legend=c("7", "8"), fill=col,  title="Grade")
legend("topright", legend=c("Male", "Female"), pch=c(15,17),  title="Gender")
```
For number of triangles, gender/grade homophily we will use ERGM. 

```{r }
net2=asNetwork(net)
summary(net2 ~  edges + triangles + nodefactor('gender')+nodematch('gender')+ nodefactor('grade')+nodematch('grade')+nodefactor('lunch')+nodematch('lunch'))
```


## ERGM model

```{r }
fit1=ergm(net2 ~ edges  + nodematch('gender') +nodematch('grade') + nodematch('lunch'))

fit2=ergm(net2 ~ edges  + nodefactor('gender')+nodematch('gender')+ nodefactor('grade')+nodematch('grade') + nodematch('lunch'))

fit3=ergm(net2 ~ edges  + nodefactor('gender')+nodematch('gender')+ nodefactor('grade')+nodematch('grade') + nodematch('lunch') +altkstar(0.5, fixed=TRUE))

#fit4=ergm(net2 ~ edges  + nodefactor('gender')+nodematch('gender')+ nodefactor('grade')+nodematch('grade') + nodematch('lunch')+gwdegree(1, fixed=TRUE))

```

```{r model-tabel, echo = FALSE, results='asis'}
texreg::knitreg(list(fit1, fit2, fit3))
```





