---
format: html
title: Sunbelt 2023 presentation
subtitle: "Differences in differences in network statistics"
embed-resources: true
execute: 
  warning: false
---

## Pairwise comparisons with the original network

We first reload the original data to take the raw values

```{r}
library(data.table)
simstats <- fread("data/02-dataprep-network-stats.csv.gz")
```

Preparing the data for the pairwise comparisons

```{r}
# Dependent variables
depvars <- c("peak_time", "peak_preval", "rt", "rt_mean", "dispersion", "gentime")

# Independent variables
indepvars <- c("inc_days", "recovery_rate", "transmission_rate")

# Listing variables starting with ergm and igraph separately
ergmvars <- grep("^ergm", names(simstats), value = TRUE)
igraphvars <- grep("^igraph", names(simstats), value = TRUE)
```

# Visualization

Plot showing the distribution of each ergmvars using an histogram

```{r}
library(ggplot2)

```



# Data processing

We now create a function to compute the average for nettype == ergm

```{r}
ergmdat <- simstats[, lapply(.SD, mean), by = .(nettype), .SDcols = c(ergmvars, igraphvars, depvars, indepvars)]

# Appending the ergm name to the columnnames
setnames(ergmdat, old = names(ergmdat), new = paste0(names(ergmdat), "_ergm"))

# Except the nettye column
setnames(ergmdat, old = "nettype_ergm", new = "nettype")

# Merging the data
simstats <- merge(
  simstats[nettype != "ergm"],
  ergmdat, by = "nettype"
  )
```

We now calculate the pairwise differences

```{r} 
# For each var in c(ergmvars, igraphvars, depvars), compute the difference
# with the variable of the same name but appended with _ergm
for (var in c(ergmvars, igraphvars, depvars, indepvars)) {
  simstats[, paste0(var, "_diff") := get(paste0(var, "_ergm")) - get(var)]
}

```

# Regression Analyses


```{r}
# For now, excluding igraph_avg_closeness_diff.
# It is usually zero b/c of degseq (has isolates)
igraphvars <- setdiff(igraphvars, "igraph_avg_closeness")

# Creating formula, one for each devpar+"_diff" using ergmvars+"diff" and igraphvars+"diff"
# as independent variables
formulas <- lapply(depvars, function(x) 
  as.formula(
    paste0(
      x, "_diff ~ I(factor(nettype)) + ",
      paste0(indepvars, "_diff", collapse = "+"), "+",
      paste0(ergmvars, "_diff", collapse = "+") , "+",
      paste0(igraphvars, "_diff", collapse = "+")
      )
      ))

# Scaling _diff variables in simstats
simstats[, paste0(ergmvars, "_diff") := lapply(.SD, scale, scale = TRUE), .SDcols = paste0(ergmvars, "_diff")]
simstats[, paste0(igraphvars, "_diff") := lapply(.SD, scale, scale = TRUE), .SDcols = paste0(igraphvars, "_diff")]
simstats[, paste0(depvars, "_diff") := lapply(.SD, scale, scale = TRUE), .SDcols = paste0(depvars, "_diff")]

# Fitting the models
models <- lapply(formulas, function(x) lm(x, data = simstats))
```



# Table 1

::: {style="font-size: 70%"}

```{r}
#| results: "asis"
# library(stargazer)
# stargazer(models, type = "text", omit.stat = c("f", "ser"), omit.table.layout = "n")
texreg::knitreg(
  setNames(
    models, 
    depvars
  ), single.row = TRUE)
```

:::